Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
more alignment wip
More polishing in case-study

In this section we sketch an existing enforcement mechanism, {\em Depth
Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
formal stack-safety property described in the previous section. (In
fact, it enforces a slightly stronger, inductive version of that property which
strengthens the local state encapsulation component by checking for
integrity and confidentiality violations at every step of the callee's
execution trace.) We validate this claim with a property-based
testing framework that can quickly uncover inconsistencies between the
enforcement mechanism and the stack-safety property.

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the {\em Depth Isolation} policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\item From callee back to caller, when execution is about to reach a
  return point.  The exit sequence retags all locations in the
  function frame with $\tagNoDepth$. At the actual return instruction,
  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
    does that now (retag unused). But I don't think it's a
    confidentiality or integrity violation.}
  \rb{Correct, and it doesn't. Some versions of the policy do clear
    the stack, retagging would be easily added.}

@@

We validated the micro-policy of the previous section by systematically
testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
extended it with a runtime monitor implementing the stack-safety
micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
 Or later?} We chose the
Coq proof assistant as the setting for our implementation for two
reasons: first, to ensure that our coinductive trace definitions are
well-formed, and reason about them\leo{Do we prove anything
  interesting anymore?}, and to leverage the power of the QuickChick
property-based testing framework~\citep{Pierce:SF4}.

@@

To write such generators we build on the work of \leo{ICFP 2013},
where they introduced {\em generation by execution} to produce
interesting programs. Generation by execution receives as an input a
partially instantiated machine state and attempts to generate an
instruction (or a sequence of instructions such as a blessed
sequence) that makes sense locally (e.g., jumps go to a potentially valid
code location and loads read from from a potentially valid stack
location). Then
we step the machine and repeat the process until we generate or
execute some target number of instructions, or if we reach a point
where the machine can't step anymore.

@@

@@

To evaluate the effectiveness of testing against our formal properties, we
used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest (here,
stack safety) to fail, and ensure that the testing framework can find
them. The bugs we used for our evaluation were either artificially generated
by us (deliberately weakening the micro-policy in ways that we expected
should break its guarantees), or actual bugs that we discovered through
testing our implementation. We elaborate on some such bugs below.

The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
  spec}. \leo{Something like: There is really no baseline to compare this
  against, but include it here for the future/reasons?}

The artificially injected bugs focus on weakening the tag propagation
and checking mechanism. \leo{Mention them all?}\bcp{Nah, especially if they
  are all listed in the table.}
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the one of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
that restriction by not checking the depth equality, or not checking
the tags at all, causing bugs of different difficulty.
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing these constraints also causes violations to the
property.

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code
%\bcp{right?}\leo{yeah}
; we re-implemented it in Coq targeting
RISC-V).  These errors ranged from trivial typos to
ones that required an
intriguingly complex setup to appear.  Notably, our first try at the
blessed call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allowed for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}

>>>>

In this section we explore how an existing enforcement mechanism, {\em
  Depth Isolation}, from \citet{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section. (In fact, it enforces a slightly stronger, inductive version
of that property which strengthens the local state encapsulation
component by checking for integrity and confidentiality violations at
every step of the callee's execution trace.) We validate this claim
with a property-based testing framework that can quickly uncover
inconsistencies between the enforcement mechanism and the stack-safety
property.

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the {\em Depth Isolation} policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
\leo{If we have the simple exposition so far, no need to mention the following sentence:}
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\item From callee back to caller, when execution is about to reach a
  return point.
  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
  At the actual return instruction,
  the $\PCname$ tag is decremented.
%  \leo{Not sure the actual policy
%    does that now (retag unused). But I don't think it's a
%    confidentiality or integrity violation.}
%  \rb{Correct, and it doesn't. Some versions of the policy do clear
%    the stack, retagging would be easily added.}

@@

We validated the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture
\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
\leo{translated from Haskell via hs-2-coq?}  and extended it with a
runtime monitor implementing the stack-safety
micro-policy.\footnote{This code will be made public. \leo{More?}} We
  chose the Coq proof assistant as the setting for our implementation
  for two reasons: first, to ensure that our coinductive trace
  definitions are well-formed, and reason about them\leo{Do we prove
    anything interesting anymore?}, and to leverage the power of the
  QuickChick property-based testing framework~\citep{Pierce:SF4}.

@@

To write such generators we build on the work of
\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, where they
introduced {\em generation by execution} to produce progams leading
abstract machines towards interesting behaviors. Generation by
execution receives as an input a partially instantiated machine state
and attempts to generate an instruction (or a sequence of instructions
such as a blessed sequence) that makes sense locally (e.g., jumps go
to a potentially valid code location and loads read from from a
potentially valid stack location). Then we step the machine and repeat
the process until we generate or execute some target number of
instructions, or if we reach a point where the machine can't step
anymore.

@@

In addition, we needed to extend this technique to handle the nested
nature of our confidentiality property: rather than generating two
initial machines that are variant of one another and letting them
execute to test for noninterference, we need to generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks the confidentiality of multiple nested calls.

@@

To ensure the effectiveness of testing against our formal properties, we
used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest (here,
stack safety) to fail, and ensure that the testing framework can find
them. The bugs we used for our evaluation were either artificially generated
by us (deliberately weakening the micro-policy in ways that we expected
should break its guarantees), or actual bugs that we discovered through
testing our implementation. We elaborate on some such bugs below.

\begin{table}[]
\centering
\begin{tabular}{c|c|c}
  Bug & Property Violated & MTTF (ms) \\
  \hline
      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:bug-table}
\end{table}

The mean-time-to-failure (MTTF) for various bugs can be found in
Table~\ref{tab:bug-table}. All experiments were run in a desktop
machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
RAM. \leo{Something like: There is really no baseline to compare this
  against, but include it here for the future/reasons?}
%
The artificially injected bugs focus on weakening the tag propagation
and checking mechanism. 
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the one of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
that restriction by not checking the depth equality (row {\em
  LOAD\_NO\_CHECK}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraints causes violations to the integrity
property (row {\em STORE\_NO\_CHECK}).

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code
%\bcp{right?}\leo{yeah}
; we re-implemented it in Coq targeting RISC-V).  These errors ranged
from trivial typos to ones that required an intriguingly complex setup
to appear.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}), in our first try the blessed call
sequence/policy combination did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured: in
particular, the smallest counterexample requires calling a function,
writing in the uninitialized location (which is always fine), calling
another function (to seal what was written for integrity purposes),
and then attempt to write in the same location again from the new
callee, which is only allowed because the enforcement still thinks
the location is unused, but violates the integrity of the caller's data.
